### 2021-03-31 referee_response1.md

We have added text to make it clearer that different arms can have different pitch angles, even if in a single arm, pitch angle may be constant with radius.

> On page 3 it is written: "A common assumption  when  measuring  galaxy pitch angle is that observed spiral arms have a constant pitch angle with radius (e.g. Davis et al. 2012;.... " This is repated a number of times in other parts of the paper. Yet this is wrong, as first duscussed by Considere and Athanassoula (who introduced the 2DFFT method) and as stressed further in Davis et al 2012.

> Davis et al say in their paper: "As pointed out by Considere &  Athanassoula (1988), this  method does  not  assume  that observed spiral structures are logarithmic. It only decomposes the observed distributions into a superposition of logarithmic spirals of different pitch angles and number of arms, which can be thought of as building blocks." The authors should read the Considere & Athanassoula paper carefully, as well as the Davis et al (2012) paper and change accordingly all parts of the paper where they make use of this assumption, or find a different way to justify it.

We thank the referee for bringing Considere & Athanassoula 1988 back to our attention. We have reviewed the paper again, and added a citation. In that paper it is very clear the method does not assume logarithmic spirals. However the first line of the abstract of Davis et al (2012) is *"A logarithmic spiral is a prominent feature appearing in a majority of observed galaxies"*, and they clearly say (Section 4.3.2) that the inclusion of non-logarithmic spirals (with significant variation of pitch angle with radius) can negatively impact measurements using 2DFFT. They also say "In general, logarithmic spirals are good approximations of the shape of galactic spiral arms". In the face of these comments, which show that the assumption of logarithmic spirals is made despite the more general applicability of the method of Considere & Athanassoula (1988), we think there is evidence in the literature that the assumption of logarithmic spirals is widespread, even if the 2DFFT method can in principle discover radial variation in phi.

@@

> Some of the assumptions made in this paper come without much/any justification. The authors should discuss the effects of each of the assumptions for which they offer no justification. E.g. is it reasonable to assume that the spread sigma_[gal] is constant? What do results from 2DFFT give? Some *quantitative* analysis of results from previous work would be useful here. If this assumption is simply based on lack of any better knowledge, the effect of this assumption should be tested numerically and discussed. But the authors could first check the literature whether there is better knowledge from previous 2DFFT work.

We have added more discussion of the assumptions which were made. The specific question about sigma_[gal] is interesting. We have added citations to work where arms in spirals are fit separately. All of them comment that pitch angles do vary between arms in spirals, but there is no discussion of how the variation changes between galaxies. In general sample sizes are small. We note this as an interesting followup question for larger samples, including potentially a larger sample run through a second phase of Galaxy Builder.

@@

> This section is rather sparse in information. However, there are 198 galaxies considered, of which 98 galaxies were shown twice to volunteers. In Lingard et al. 2020, the Galaxy Builder paper - GBpaper for short - the stellar masses are shown in Figure 4. For 120 galaxies the stellar mass log M_* is between 9.45 and 10.03, and there after there are fewer galaxies up until the upper mass range of 11.0. Please provide a Figure with the histogram of stellar masses of the galaxies studied here. This sample is skewed towards galaxies with lower stellar mass: a quick analysis of the galaxies in Kennicutt (1981) which are also in S4G shows very few galaxies in the range of log M_* between 9.45 and 10.03. Please indicate what the consequences are of this selection effect for the general conclusions.

We have added a histogram (which was previously published in Lingard et al. 2020) as a new Figure 3 in Section 2.2. We have also added a histogram, and an investigation of how pitch angle varies with stellar mass in Section 3.1 in response to another request below.

> Since a sample of 98 galaxies were shown twice to (different?) volunteers, it is of interest to compare the results for the same galaxies: did the resulting pitch angles, number of arms, etc. all agree ? One of the data files available on the GitHub website indicates that this is not the case, but that website is poorly documented.

 We added a line to point the reader to Lingard et al. (2020) where this is discussed, and also add a comment that in this set of 98 galaxies, we find that 93% have the number of arms the same within +/-1 (53% have exactly the same number of arms).

@@

We agree. Our statistical analysis incorporates this (which is one of the reasons we don't find anything significant). We have made this clearer in the text.

> 3. Section 3.1

> You report a final sample of 139 galaxies with 261 spiral arms, i.e. less than 2 arms per galaxy. Please give details of why this is so. Please also list how many of the galaxies are multiarmed, and state the number of arms.

When we looked into this we found this needed to be updated to 129 galaxies with 247 spiral arms which is closer to two arms per galaxy (1.9). We are unsure of the reason for this discrepancy. All plots and numberical result have been updated.

@@

* 159 total galaxies, 129 of which with identified spiral arms
* 247 total spiral arms
* 238,433 total points
  *  30  galaxies had 0 arm(s)
  *  38  galaxies had 1 arm(s)
  *  68  galaxies had 2 arm(s)
  *  19  galaxies had 3 arm(s)
  *  4   galaxies had 4 arm(s)

This has been entered into the text.

> In Figure 4 I count about 40 1-armed spirals, 22 3-armed spirals, 14 4-armed spirals, and about 63 2-armed spirals (more difficult to count). Why are there so many 1-armed spirals, and why for these, for pitch angles below about 14 degrees, is PHI_arm not equal to PHI_gal ?

 We have found that in some cases, while most volunteers drew two arms, the clustering/aggregation code was unable to cluster them, so the final method reported a single arm. This happened in 35 cases (the majority of the 1 arm galaxies), and is a consequence of clustering noisy data. Where arms are detected the pitch angles are reliable, and our method of calculating galaxy averages accounts for this possible bias of error caused by missing arms. However the presence of this fraction of single-identified spiral arm galaxies drives the average number of arms per galaxy down below the real value, and because the most common number of actual arms is two, it is particularly notable in resulting in reports of one-armed galaxies. We have added some text explaining this, and cautioning against over interpreting our reports on the number of arms. We reiterate that this should not significantly impact the results on pitch angle.

@@

> 4. Section 3.2.1

> I quote: "Morphological classification commonly links bulge size to spiral tightness, and such a link is implied by the Hubble Sequence (Sandage 2005, Gadotti 2009, Buta 2013). Some studies have indeed reported a link between measured spiral galaxy pitch angle and bulge size (i.e. Hart et al. 2017, Davis et al. 2019), while others have not found any significant correlation (Masters et al. 2019).

> Now there is a large overlap in authors between Hart et al. (2017) and Masters et al. (2019): 7 people, 5 of which are also co-authors of this paper. Surely these 5 people should be able to explain WHY they found no link between pitch angle and bulge size in one paper, and found a link in another one ??

We think addressing this, while interesting, is beyond the scope of this paper. The Hart et al. (2017) result really showed that pitch angle depended on disc mass more than bulge size (but more massive discs tend to have larger bulges). In the interest of keeping to the point of this paper, we have removed that citation, but also note that these correlation results can depend on both how bulge size and pitch angles are measured.

@@

This error has been corrected.

> The range of values of PHI_gal is from 8 to about 50 degrees, and there is no indication that the lower limit in the pitch angle is poorly determined. Note that Kennicutt (1981) find a lower limit of 0 degrees. Setting the lower limit to 15 degrees, so as just to be able to conclude that a uniform cot PHI_gal or cot PHI_arm distribution cannot be excluded is thus completely arbitrary. It would be best to display in a second panel the distribution of cot PHI_gal or cot PHI_arm with a lower limit of 8 degrees, so as to alert the reader to the strong influence of this assumption.

 The issue here is that the model itself includes an arbitrary lower pitch angle - Pringle & Dobbs (2019) make no attempt to give any physical motivation for what that minimum angle should be, and in their test against data chose 11.5 degrees as the limit - largely because that's where inspection of the distribution of cot(pitch angle) in the sample they investigate deviates from constant.

 We think the referee makes an excellent point that the previous explanation was a little confusing. What we have really determined in our study here is that the cot(phi) uniform "model" fits the data well when the minimum pitch angle for winding spirals is around 15 degrees. We also determine that if the minimum pitch angle is 10 degrees the model cannot fit our data. This suggests that a minimum pitch angle for this model must be above 10 degrees,(e.g. 15 degrees). That's a specific observation that a more detailed model based on the simple model of Pringle & Dobbs (2019) would have to explain. We have made edits to the text to make this point, which we think is a really nice addition to the paper.

@@

> I quote: "We do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory (see Section 3.2.2). Just in case you reach a conclusion in that section, and want to mention it in this conclusion section, could you please formulate this in proper English? With so many native English/American speakers on the list of authors this ought not to be too much asked, and would help the reader understand better, as well as save considerable work for the copy editor.

We have reworded this section.

> Another quote: "In this work, we assume that spiral arms are equally likely to be identified and recovered at all pitch angles." OK, but then you should NOT arbitrarely discard low pitch angle values when testing the Pringle & Dobbs 2019 hypothesis.

We assume no selection bias in Galaxy Builder results, which is not necessarily valid, however the requirement for a lower limit of pitch angle is needed for the test and we have added to the explanation of why in the text (also see comments above).

@@

> 8. Data availibility

> The GitHub references refer to poorly documented working directories, and an interested researcher needs to make a considerable effort to start finding the basic information. It would be much more helpful to provide the most relevant information about the only several 100s of galaxies in machine readable tables, which can be stored at e.g. the CDS.

We apologise, we mistakenly linked a working respository, rather than the planned public one, which is available at https://github.com/tingard/hierarchical-modelling-of-spiral-pitch-angle.

Following this feedback, we have additionally added (see at end of the README) a description of the contents of each of the Python Notebooks.

### 2021-03-31 referee_response2.md

# Referee Response, second round.

> Major comments:

> 1. In the Introduction, the authors change the three criteria of the ordering in the Hubble classification sequence. However, Sandage (2005) repeats these several times: the size of the central bulge, the degree of resolution into condensations, and the openness of the spiral arms.

> The authors change the "degree of resolution into condensations" to "the degree of resolution of the spiral arms (how 'patchy' they are)". Yet the word patchy is never used in this context, and for flocculent galaxies (where there are no well described spiral arms), such as NGC 2841 (type SAT3 in the Revised Shapley Ames Catalog), an ordering is nonetheless given, in coherence with all three criteria.

> Please refrain from reformulating the Hubble classification to a scheme based on your own criteria: e.g. the prototype flocculent spiral galaxy NGC 2841 has condensations in its disc, but no clear spiral arms.

OK. We have made edits to use the recommended language from Sandage (2005) and refrained from adding any context to help modern readers (I know this is snark - which I will edit later).

> 2. In this context, it is striking that the authors now choose NGC 4689 as a example of a flocculent galaxy, even though spiral structure is visible, in particular the eastern arm. True, Elmegreen & Elmegreen (1987) list it as a class 3 spiral, i.e. in the same category as NGC 2841, but the spiral arms are clearly there in the 3.6 micron Spitzer image. I suggest using NGC 2841 as a good example for a flocculent galaxy, and note that for NGC 5055 there is a paper by Michele Thornley -  1996ApJ...469L..45T - pointing out that in the K'-band there is spiral structure seen, even though it is also listed as a prototype flocculent spiral in Elmegreen & Elmegreen (1987).

We have edited to use NGC 2841 as the example of a flocculent galaxy ** TO DO - I put the SDSS image of it in the drive.  **

> 3. The authors write on page 2 left side line 38: Spiral tightness is also observed to correlate with rotation curve shape (Seigar et al. 2005), with rising rotation curves creating more open spiral structure.

> I think you mean that galaxies with rising rotation curves have more open spiral structure. I don't think rotation curves are creating anything. More likely, the associated lower shear allows more open spirals.

We have edited this sentence as suggested.

> 4. In section 3.2.2 the authors write: One of the predictions of Manifold theory is that pitch angle increases with bar strength (Athanassoula et al. 2009b).

> Please make this section self-contained: ... pitch angle increases with bar strength (Athanassoula et al. 2009b), as evaluated at the Lagragian L_1 and L_2 points. The bar strength evaluated that manner is very specific, and in Athanassoula et al. 2009b it is shown that a correlation with Q_bar is much less strong. Since you use here p_bar as a proxy for bar strength, it is only remotely correlated with the bar strength defined in Athanassoula et al. 2009b. True, you admit that your test is not very convincing, but the interested reader should not have to consult explicitly the Athanassoula et al. 2009b paper to find out why.

We have edited this section as suggested.

> 5. In the conclusion section, the formulation of the conclusions on section 3.2.2. are much stronger than justified: you say that you cannot completely rule out the manifold theory, even though you admit that the proxy you use is not convincing. Since hurried readers could read this conclusion section without reading the bulk of the paper, you leave them with the impression that you almost rule out the Manifold theory, while this is not the case, since your proxy for bar strength is very poor, and your sample is small. Please harmonize your conclusions with the reality you describe in the main section.

> 6. In the conclusions, page 11 left side lines 48-49, you suggest that the absence of galaxies with low pitch angles is not a selection effect. How does this square with the data of Kennicutt (1981), where nearly all Sa galaxies, as classified in the RSA catalogue, have pitch angles less than 10 degrees? Are you sure that you are not missing the tightly wound early type spiral galaxies?

> I did a cursory inspection of the early type spirals in the Yu & Ho (2018) sample, and find that most of the early type spirals (S0/a, Sa and Sab) are actually barred spirals when checking the classification in Hyperleda and displaying them in the legacysurvey viewer. Perhaps the authors can check the distribution of the results of Kennicutt (1981) in a diagram such as Figure 9, to see whether his sample is in agreement with the ideas of Pringle & Dobbs (2019)?

> Minor comments and typos:

Thank you for noting these minor typos and other errors. We have corrected these, and some other ones we noted in editing.

> 1. page 2 right side line 50: disk - please use English spelling throughout: disc.

> 2. page 3 left side line 40: we further separate the galaxies ...

> 3. page 4 left side line 21: and report the pitch angle

> 4. page 4 left side line 53: which results in a sample favouring (3 times "which" is too much)

> 5. page 4 left side line 60: of these classifications

> 6. page 6 right side line 19: morphological

> 7. page 10 right side line 27: you write "pitch angle calculation; obtaining" but the ; should disappear

### 2021-03-31 sections/abstract.tex
Starting edits for 2nd report.

% !TEX root = ../main.tex
Spiral structure is ubiquitous in the Universe, and the pitch angle of arms in spiral galaxies provide an important observable in efforts to discriminate between different mechanisms of spiral arm formation and evolution. In this paper, we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project. We present a new approach to deal with the large variations in pitch angle between different arms in a single galaxy, which obtains full posterior distributions on parameters. We make use of our pitch angles to examine previously reported links between bulge and bar strength and pitch angle, finding no correlation in our data {\bf (with a caveat that we use observational proxies for both bulge size and bar strength which differ from other work)}. We test a recent model for spiral arm winding, which predicts uniformity of the cotangent of pitch angle between some unknown upper and lower limits, {\bf finding our observations are consistent with} this model of transient and recurrent spiral pitch angle {\bf as long as the pitch angle at which most winding spirals dissipate or disappear is larger than 10$^\circ$.}

>>>>

% !TEX root = ../main.tex
Spiral structure is ubiquitous in the Universe, and the pitch angle of arms in spiral galaxies provide an important observable in efforts to discriminate between different mechanisms of spiral arm formation and evolution. In this paper, we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project. We present a new approach to deal with the large variations in pitch angle between different arms in a single galaxy, which obtains full posterior distributions on parameters. We make use of our pitch angles to examine previously reported links between bulge and bar strength and pitch angle, finding no correlation in our data  (with a caveat that we use observational proxies for both bulge size and bar strength which differ from other work). We test a recent model for spiral arm winding, which predicts uniformity of the cotangent of pitch angle between some unknown upper and lower limits, finding our observations are consistent with this model of transient and recurrent spiral pitch angle as long as the pitch angle at which most winding spirals dissipate or disappear is larger than 10$^\circ$.

### 2021-03-31 sections/introduction.tex
Starting edits for 2nd report.

Spiral structure is present in a majority of massive galaxies (e.g. \citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}) yet the formation mechanisms through which spiral structure originates are still hotly debated (e.g. \citealt{2014PASA...31...35D}). Spirals are as diverse as the theories proposed to govern their evolution, from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral
\sout{. These variations on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively}
(\citealt{2011ApJ...737...32E}; examples of each type are shown in Figure \ref{fig:spiral-galaxy-types}). The Hubble classification scheme \citep{1926ApJ....64..321H} and its revisions and expansions \citep{1961hag..book.....S,1991rc3..book.....D} contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by \textbf{the degree of resolution of the spiral arms (how ``patchy” they are)}, how tightly wound they are and the prominence of a central bulge. \textbf{Building on this, \citet{1982MNRAS.201.1021E} found that flocculent spirals are more prevalent in unbarred, isolated galaxies. The presence of a bar, a binary companion or group membership result in a higher fraction of observed grand design spiral patterns.}

\begin{figure*}
  \includegraphics[width=15cm]{plots/galaxy_types.jpg}
  \caption{Examples of the different types of spiral galaxy present in the sky. The left column shows the \textbf{grand design spiral NGC 5248}, the middle shows the \textbf{many-armed spiral NGC-3184} and the right shows the \textbf{flocculent spiral NGC-4689}. Images were taken with the Sloan Digital Sky Survey Telescope.}
  \label{fig:spiral-galaxy-types}
\end{figure*}

\textbf{
Whatever kind of spiral is present in a disc galaxy, there is plenty of evidence to suggest that they have a significant role on the overall evolution of that galaxy. For example, a majority of the population of young stars in a galaxy are located in its spiral arms \citep{2011EAS....51...19E}, and there is evidence that spiral arms may trigger star formation \citep{2013A&A...560A..59C} perhaps via their ability to promote the growth of Giant Molecular Clouds \citep{2014IAUS..298..221D}. The rearrangement of disc gas and stars driven by spiral arms (e.g. \citealt{2018MNRAS.476.1561D}) may lead to the formation of disc-like bulges (commonly called ``pseudobulges''; e.g. \citealt{2004ARA&A..42..603K}), which are prevalent in most spiral galaxies, including those without bars \citep{2010ApJ...716..942F}. Studies of spiral morphology have also found interesting correlations with other galactic properties, such as a correlation between the tightness of spiral arms and central mass concentration (\citealt{2019ApJ...871..194Y}, though neither \citealt{2017MNRAS.472.2263H} nor \citealt{2019MNRAS.487.1808M} found such a relation in large samples). Spiral tightness is also observed to correlate with with rotation curve shape (\citealt{2005MNRAS.359.1065S}), with rising rotation curves creating more open spiral structure. These predictions and observations provide compelling reasons for continued investigation of the underlying rules and dynamics of spiral structure, as doing so is essential for understanding the secular evolution of disc galaxies.
}

@@

One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanged for many rotational periods \citep{1964ApJ...140..646L}. The motivation for static waves with small numbers of arms \sout{(with a preference for $m=2$)} was primarily observational: most disc galaxies observed at the time showed spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt. \textbf{This, in combination with theoretical arguments about the ``winding problem", motivated the original static density waves of \citet{1964ApJ...140..646L}, to which swing amplification was added by \citet{Toomre1981}  to provide a way to counteract the short lifetime of stellar density waves.}

@@

Many other galactic components may correlate with spiral morphology, including bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2008ApJ...678L..93S}, \citealt{2017MNRAS.471.2187D}, \citealt{2019MS&E..571a2118A}). Larger bulges and more massive central black holes have both \textbf{been observed to} correlate with more tightly wound spiral arms. \textbf{We can also test this with the data presented in this paper.}

\textbf{This paper is structured as follows: In Section \ref{section:method} we introduce methods to measure galaxy pitch angle, and present our sample, and our Bayseian hierarchical modeling method making use of \textit{Galaxy Builder} to estimate galaxy and population pitch angles. Section \ref{section:constraints-on-galaxy-phi} presents our general constraints on pitch angles in our section, while} Section \ref{section:morphology-comparison} examines the correlation between pitch angle and bulge size implied by the Hubble sequence, and pitch angle and bar strength implied by Manifold theory. Section \ref{section:spiral_winding} investigates spiral arm winding using the test derived by \cite{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$). \textbf{We provide a summary and conclusions in Section \ref{section:summary}.}  Where necessary, we make use of $H_0 = 70\ \text{km}\ \text{s}^{-1}\ \text{Mpc}^{-1}$.

>>>>

Spiral structure is present in a majority of massive galaxies (e.g. \citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}) yet the formation mechanisms through which spiral structure originates are still hotly debated (e.g. \citealt{2014PASA...31...35D}). Spirals are as diverse as the theories proposed to govern their evolution, from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral.
(\citealt{2011ApJ...737...32E}; examples of each type are shown in Figure \ref{fig:spiral-galaxy-types}). The Hubble classification scheme \citep{1926ApJ....64..321H} and its revisions and expansions \citep{1961hag..book.....S,1991rc3..book.....D} contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by the \textbf{openness of the spiral arms, the} degree of resolution \textbf{into condensations} and the prominence of a central bulge. Building on this, \citet{1982MNRAS.201.1021E} found that flocculent spirals are more prevalent in unbarred, isolated galaxies. The presence of a bar, a binary companion or group membership result in a higher fraction of observed grand design spiral patterns.

\begin{figure*}
  \includegraphics[width=15cm]{plots/galaxy_types.jpg}
  \caption{Examples of the different types of spiral galaxy present in the sky. The left column shows the grand design spiral NGC 5248. the middle shows the many-armed spiral NGC-3184 and the right shows the flocculent spiral NGC {\bf 2841}. Images were taken with the Sloan Digital Sky Survey Telescope.}
  \label{fig:spiral-galaxy-types}
\end{figure*}

Whatever kind of spiral is present in a disc galaxy, there is plenty of evidence to suggest that they have a significant role on the overall evolution of that galaxy. For example, a majority of the population of young stars in a galaxy are located in its spiral arms \citep{2011EAS....51...19E}, and there is evidence that spiral arms may trigger star formation \citep{2013A&A...560A..59C} perhaps via their ability to promote the growth of Giant Molecular Clouds \citep{2014IAUS..298..221D}. The rearrangement of disc gas and stars driven by spiral arms (e.g. \citealt{2018MNRAS.476.1561D}) may lead to the formation of disc-like bulges (commonly called ``pseudobulges''; e.g. \citealt{2004ARA&A..42..603K}), which are prevalent in most spiral galaxies, including those without bars \citep{2010ApJ...716..942F}. Studies of spiral morphology have also found interesting correlations with other galactic properties, such as a correlation between the tightness of spiral arms and central mass concentration (\citealt{2019ApJ...871..194Y}, though neither \citealt{2017MNRAS.472.2263H} nor \citealt{2019MNRAS.487.1808M} found such a relation in large samples). Spiral tightness is also observed to correlate with with rotation curve shape (\citealt{2005MNRAS.359.1065S}), with \textbf{galaxies with rising rotation curves having} more open spiral structure. These predictions and observations provide compelling reasons for continued investigation of the underlying rules and dynamics of spiral structure, as doing so is essential for understanding the secular evolution of disc galaxies.

@@

One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanged for many rotational periods \citep{1964ApJ...140..646L}. The motivation for static waves with small numbers of arms was primarily observational: most disc galaxies observed at the time showed spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt. This, in combination with theoretical arguments about the ``winding problem", motivated the original static density waves of \citet{1964ApJ...140..646L}, to which swing amplification was added by \citet{Toomre1981}  to provide a way to counteract the short lifetime of stellar density waves.

@@

Many other galactic components may correlate with spiral morphology, including bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2008ApJ...678L..93S}, \citealt{2017MNRAS.471.2187D}, \citealt{2019MS&E..571a2118A}). Larger bulges and more massive central black holes have both been observed to correlate with more tightly wound spiral arms. We can also test this with the data presented in this paper.

This paper is structured as follows: In Section \ref{section:method} we introduce methods to measure galaxy pitch angle, and present our sample, and our Bayseian hierarchical modeling method making use of \textit{Galaxy Builder} to estimate galaxy and population pitch angles. Section \ref{section:constraints-on-galaxy-phi} presents our general constraints on pitch angles in our section, while Section \ref{section:morphology-comparison} examines the correlation between pitch angle and bulge size implied by the Hubble sequence, and pitch angle and bar strength implied by Manifold theory. Section \ref{section:spiral_winding} investigates spiral arm winding using the test derived by \cite{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$). We provide a summary and conclusions in Section \ref{section:summary}.  Where necessary, we make use of $H_0 = 70\ \text{km}\ \text{s}^{-1}\ \text{Mpc}^{-1}$.
